Trauma treatment delay

### 2023-12-11 manuscriptlongversion.Rmd
text

@@

Misclassification bias as the conference may reach a consensus that is
wrong, alternatively that the two specialized nurses going through the
patients flagged for audit filters might missclassify patients or that
the audit filter misses a patient with ofi.

@@

The analysis shows that some variables seem to have a correlation with
patients receiving delayed treatment. ISS and respiratory rate both have
a trend that patient with higher values seem to have a lower incidence
of delayed treatment, both unadjusted (ISS, OR 0.96, 95% CI
0.95-0.98)(RR, OR 0.95, 95% CI 0.91-0.98) and adjusted (ISS, OR 0.97,
95% CI 0.95-0.99)(RR, OR 0.96, 95% CI 0.93-0.99). GCS had the opposite
correlation, with lower GCS often having lower incidence of delayed
treatment, both unadjusted (OR 1.13, 95% CI 1.07-1.20) and adjusted (OR
1.08, 95% CI 1.00-1.15). Number of injuries had the correlation of
higher number of injuries having lower incidence of delayed treatment
unadjusted (OR 0.94, 95% CI 0.91-0.98), but that correlation was not
seen when adjusted for. The three remaining variables, age, gender and
SBP, showed no correlation with the observed outcome.

@@

These results show that there seems to be a correlation between some
variables and the presence of delayed treatment. A higher GCS had a
correlation with delayed treatment while a lower ISS had the same
correlation. This suggests that a less injured and more awake patient
has a higher risk of receiving delayed treatment. Compared to previous a
previous study analysing the same data but with ofi as the outcome, this
result is surprising as that study came to the opposite result
[@Pat_factors]. This can possibly be explained by that the study looked
at all ofi, whereas this study concentrates on delayed treatment.
Possible explanations to the correlation include the idea that
unconsious and more severely injured patients receive more immediate
attention at the hospital, leading to fewer delays.

@@

Another limitation with this study is that logistical regressions
assumes that the data is binomial, which the data used is not. The
analysis assumes a linear correlation between the dependent variable and
the independent variable, which is best analysed in a normal
distribution. However, Fischers test showed an association between 4 of
the 7 variables, all of them being the ones who showed an increased or
decreased risk of delayed treatment, showing that these results still
are significant.

@@

Further discussing the results and discussing it's clinical
applications, the results can be interpreted as that it is mostly
uninjured and awake patients that receive delayed treatment. The
clinical applications for this are not clear, but might suggest that
personnel working with trauma patients should be aware of that people
that seem uninjured might be more likely to receive delayed treatment
and trying to lower that risk.

@@

In future studies building upon this one, various methods to further
examine the correlation between delayed treatment and various variables,
alternatively using another ofi as the outcome variable. Examining
different variables such as highest hospital care lever, intubation
status or first emergency procedure performed might be interesting
variables, among others. To be able to do this, more patients in the
outcome group would be needed, either by expanding the register to
several hospitals, or expanding the outcome group by adding different
ofi. This however would make the study less specific for delayed
treatment. Further studies might also look into different ofi other than
delayed treatment, such as

@@

>>>>

*Introduction:* Trauma remains a global health challenge, leading to
millions of deaths each year. The biggest impact can be seen in low to
middle income countries, but in high income countries, trauma systems
have been developed to lower the mortality and morbidity associated with
trauma. To further improve on these systems, conferences are held to
better the care patients receive and to identify errors made, called
opportunities for improvement, e.g. delayed treatment. *Aim:* This
retrospective cohort study aims to identify patient level factors that
are linked with receiving delayed treatment among trauma patients.
*Material and Methods:* This study uses the database of Karolinska
university hospital, the equivalent of a level 1 trauma center, to run a
logistical regression linking the outcome, delayed treatment, to
variables, observed patient level factors. Patients treated for trauma
at the hospital between 2012 and 2022 were included, with ISS\>9 and/or
trauma code activation. Exclusion criterias were not being screened for
ofi and missing values in the used variables, age, gender, ISS, SBP, RR,
GCS and number of injuries. *Results:* Higher ISS and RR was linked to
higher rates of delay in treatment (ISS, OR 1.03, 95% CI 1.01-1.06)(RR,
OR 1.04, 95% CI 1.01-1.08), while a lower GCS also was linked to delayed
treatment (OR 0.93, 95% CI 0.87-1.00). Number of injuries showed a
correlation to higher rates of delay, but not when adjusted for. The
remaining variables showed no correlation. *Conclusion:* Our study
suggests that it is mostly more severely injured and less awake patients
that receive delayed treatment.

*Inledning:* Trauma är fortfarande en global hälsoutmaning som skördar
miljontals liv varje år. Den största effekten kan ses i låg till
medelinkomstländer, men i höginkomstländer har traumasystem utvecklats
för att sänka morbiditeten och mortaliteten relaterat till trauma. För
att ytterliggare förbättra dessa system hålls konferenser för att
förbättra vården patienter får och identifiera misstag som gjorts, så
kallade förbättringsmöjligheter, t.ex. försenad vård. *Syfte:* Denna
retrospektiva kohortstudie syftar till att identifiera patientfaktorer
kopplade till försenad vård bland traumapatienter. *Material och Metod:*
Denna studie använder Karolinska universitetssjukhusets databas,
motsvarande ett traumacenter på nivå 1, för att köra en logistisk
regression som kopplar utfallet, försenad vård, till variabler,
observerade patientnivåfaktorer. Patienter som behandlades för trauma på
sjukhuset mellan 2012 och 2022 inkluderades, med ISS\>9 och/eller
traumakodaktivering. Exklusionskriterier var om patient inte screenades
för ofi samt saknade värden i de använda variablerna, ålder, kön, ISS,
SBT, AF, GCS och antal skador. *Resultat:* Högre ISS och AF var kopplade
till högre grad av försenad vård (ISS, OR 1.03, 95% CI 1.01-1.06)(RR, OR
1.04, 95% CI 1.01-1.08), medan ett lägre GCS också var kopplad till
försenad vård (OR 0.93, 95% CI 0.87-1.00). Antalet skador visade först
ett samband med högre förseningsfrekvens, men inte i den justerade
analysen. De återstående variablerna visade ingen korrelation.
*Slutsats*: Vår studie tyder på att det mestadels är svårare skadade och
mindre vakna patienter som får försenad behandling.

@@

Misclassification bias as the conference may reach a consensus that is
wrong, alternatively that the two specialized nurses going through the
patients flagged for audit filters might misclassify patients or that
the audit filter misses a patient with ofi.

@@

The analysis shows that some variables seem to have a correlation with
patients receiving delayed treatment. ISS and respiratory rate both have
a trend that patient with lower values seem to have a lower incidence of
delayed treatment, both unadjusted (ISS, OR 1.04, 95% CI 1.02-1.05)(RR,
OR 1.06, 95% CI 1.02-1.09) and adjusted (ISS, OR 1.03, 95% CI
1.01-1.06)(RR, OR 1.04, 95% CI 1.01-1.08). GCS had the opposite
correlation, with higher GCS often having lower incidence of delayed
treatment, both unadjusted (OR 0.88, 95% CI 0.83-0.94) and adjusted (OR
0.93, 95% CI 0.87-1.00). Number of injuries had the correlation of
higher number of injuries having higher incidence of delayed treatment
unadjusted (OR 1.06, 95% CI 1.02-1.10), but that correlation was not
seen when adjusted for. The three remaining variables, age, gender and
SBP, showed no correlation with the observed outcome.

@@

These results show that there seems to be a correlation between some
variables and the presence of delayed treatment. A lower GCS had a
correlation with delayed treatment while a higher ISS had the same
correlation. This suggests that a more heavily injured and less awake
patient is more prone to receiving delayed treatment. Compared to
previous a previous study analysing the same data but with ofi as the
outcome, this result is similar to what they found using the same
database [@Pat_factors]. As the mentioned study suggests, this could be
explained by that lower GCS indicates severe trauma, aswell as higher
ISS. More injured patients often entails a more complex management and
require more interventions. More interventions leave more room for delay
to these interventions, why they are more prone to delay in treatment.

@@

Another limitation with this study is that logistical regressions
assumes that the data is binomial, which the data used is not. The
analysis assumes a linear correlation between the dependent variable and
the independent variable, which is best analysed in a normal
distribution. However, Fischers test showed a significant difference
between the groups in 4 of the 7 variables, all of them being the ones
who showed an increased or decreased risk of delayed treatment, showing
that these results still are significant.

@@

Further discussing the results and discussing it's clinical
applications, the results can be interpreted as that it is mostly
injured and less awake patients that receive delayed. The clinical
applications for this are not clear, but caregivers should already be
aware that the more injuries a patients has, the more room for error
exists. It is good though to receive a reminder that it's important to
always work towards minimizing the delay for these injured patients.

@@

In future studies building upon this one, various methods to further
examine the correlation between delayed treatment and various variables,
alternatively using another ofi as the outcome variable. Examining
different variables such as highest hospital care lever, intubation
status or first emergency procedure performed might be interesting
variables, among others. To be able to do this, more patients in the
outcome group would be needed, either by expanding the register to
several hospitals, or expanding the outcome group by adding different
ofi. This however would make the study less specific for delayed
treatment. Further studies might also look into different ofi other than
delayed treatment, such as missed injury, faulty communication or
triage.

@@

In conclusion, based on this small population, the results suggests that
it is mostly more injured and less awake trauma patients that receive
delayed treatment. However, due to the limitations of this study, the
result should only be seen as something to keep in mind when treating
patients as further studies need to be conducted on this subject.

#
